Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-10-03 paper.Rnw
couple of small edits

We apply LIME using six sets of tuning parameter values to all observations in the \data \ test set to observe variability across tuning parameter values. A quantile bin based simulation method (samples are simulated uniformly from a specified number of quantile bins) is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, because the complexity of the explanations increases with the number of bins. Note that 4 quantile bins are the default method in \emph{lime}. The sixth application of LIME uses a kernel density simulation method (samples are drawn from kernel density approximations of the feature distributions). The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. (See Appendix~\ref{lime-details} for descriptions of the tuning parameters.)

@@

Before introducing the visual diagnostics, let us consider a commonly used visualization of a LIME explanation that is shown in Figure \ref{fig:figure-04} (left). The plot is created using the \emph{lime} R package and provides a visual representation of a LIME explanation for the prediction of interest indicated in \autoref{fig:figure-03}. In particular, the explanation is obtained using 4 quantile bins to simulate the data. In this scenario, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the explainer model features.

@@

Based on the plot on the left of \autoref{fig:figure-04} alone, it is not possible to make an informed assessment of the explanation. For a further assessment of the explainer model, we turn to an \emph{explanation scatterplot}: a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method of four quantile bins.

@@

First, consider the claim of localness. The weights decay relatively slowly  outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bins containing the prediction of interest. However, it is  difficult to say if the claim of localness has been violated, because a definition of a local region is not specified. Depending on what region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless, it is possible to say that the weights assigned to the simulated data do not capture  the local region identified by the intersection of the quantile bins.

Now, consider the claim of faithfulness: It can be said that the majority of the points in the $x_2$ quantile support a prediction of blue, which is captured by the bar supporting a prediction of blue in the LIME explanation, and a similar statement can be made about the $x_1$ quantile bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest performs well at capturing the sine curve classification boundary by creating various sized rectangles consisting of predictions with similar probabilities, which the LIME explanation does not pick up on. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.

@@

\autoref{fig:figure-06} shows a feature heatmap for the LIME applications to the 100 observations in the \data \ test set. The most important and second most important features selected by lime are shown in in the top and bottom facets, respectively. For the quantile bins, the original features prior to the indicator variable transformation are included since it is obvious that different features would be selected when the sizes of the bins change. This figure shows that for kernel density and 2 quantile bins, LIME selects $x_1$ as the most important feature and $x_2$ as the second most important feature across all cases in the test set. These explanations are not local. There is variability in the features selected by LIME for 3 to 6 quantile bins suggesting more local explanations. There are signs of vertical striping, which suggests a dependence on tuning parameters. Note that the explanations from 3 and 5 quantile bins include the selection of the random noise variable ($x_3$) as an important variable in many predictions, which should not be the case. The pattern seen in the explanations for 3 to 6 quantile bins may suggest local explanations, but the dependence on tuning parameters makes it unclear which set of explanations to use.

@@

All three metrics suggest that the kernel density method performed well, but the metrics disagree for the quantile bins methods. Average $R^2$ and average fidelity rank the performance of the number of quantile bins the same (2 quantile bins perform the best and 6 quantile bins perform the worst). In fact, these two metrics appear to have a mirrored  relationship in this example. MSEE provides almost the exact opposite results with  6 quantile bins performing the best and 2 quantile bins performing the worst. Average fidelity and MSEE are similar metrics, so it is not surprising that they would agree in this example. Another factor might be that MSEE only takes the prediction of interest  into account and not the full simulated dataset. The contradiction between metrics makes it difficult to identify which simulation method to trust.

@@

In response, \citet{hare:2017} proposed an automated machine learning method for bullet matching to complement a visual inspection by firearm examiners. Based on high-resolution topological scans of land engraved areas, \citet{hare:2017} obtain signatures of striations from two bullet lands (\autoref{fig:figure-09}). Nine features quantifying the similarity of signatures, such as the cross-correlation function, the distance between signatures, and the number of matching striae, are extracted and used to train a random forest \kge{\sout{(available in the \emph{bulletxtrctr} R package)}} to determine the probability of a comparison resulting from the same source (matching signatures) or from different sources (non-matching signatures). The model \kge{in \citet{hare:2017}} was trained on a set of scans of bullets from the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}, which included \kge{10,384 land-to-land comparisons \citet{hare:2017}}. \kgc{Moved this sentence to the end of the paragraph:} See the supporting information for additional information on the \kge{signature similarity features\sout{data}}.

@@

\kge{We train a new random forest model using \Sexpr{format(dim(bullet_train)[1], big.mark = ",")} land-to-land comparisons from bullets in the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009} using the same nine comparison features used by \citet{hare:2017} to predict whether the two lands are a match.} We apply the trained random forest model to another set of bullets from the Hamby study with \Sexpr{dim(bullet_test)[1]} rows of land comparisons. \kge{See the supporting information for further descriptions of the data.}

We first consider a global visualization of the relationship between the random forest predictions and the model features with a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) \kge{\sout{predictions}} (\autoref{fig:figure-10}). \kge{\sout{The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.}} \kge{The training data observations show a clear difference in feature values based on whether the corresponding random forest probability is close to 1 or 0. For example, observations with a random forest probability close to 1 tend to have high values of rough correlation, cross correlation function, and matches and low values of mismatches, distance, and non-consecutivley matching striae. The observations in the test data that are classified incorrectly by the random forest tend to have feature values similar to observations in the training data with similar random forest scores. For example, the observations in the test set known to be a match but have a random forest score below 0.5 have relatively low values of rough correlation, cross correlation function, and matches and relatively large values of mismatches, distance, and non-consecutivley matching striae compared to observations with a random forest probability close to 1.}

LIME is applied to all test set observations using different tuning parameter values: 12 sampling methods (2-6 equally spaced bins, 2-6 quantile bins, kernel density estimation, and normal approximation) and 3 Gower exponents (0.5, 1, and 10). Thus, a total of $12\times 3=36$ different applications of LIME are performed. We specify that each LIME explanation return 3 features and feature selection is performed by selecting the features with the highest weights in a ridge regression model (default option).

@@

First, \kge{2 quantile bin \sout{the density}} simulation method \kge{applications} produce the same explanations for almost all cases and LIME tuning parameter values suggesting the LIME explanations are global and not local. Second, within a \kge{simulation method \sout{bin based simulation method}}, the features selected by LIME for an observation \kge{\sout{often vary by the number of bins but}} do not appear to vary by the Gower power. \kge{However, the LIME explanation for an observation often varies across simuluation methods.} With the equal bins, there are vertical stripes that suggest a dependence of the LIME explanations on the number of bins. The vertical stripes are not as apparent with the quantile bins. Lastly, there are clear differences between the LIME explanations produced by the bin based simulation methods for the matches and non-matches. This suggests that the features the random forest uses to classify a match or non-match are different.

@@

@
\caption{Assessment metric plot for the applications of LIME on the bullet comparison data test set. \kge{\sout{The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.} There are discrepancies in performance across metrics for a set of tuning parameter values, but based on the overall performance of all three metrics, 4 equal bins with a Gower power of 0.5 appears to perform the best.}}
\label{fig:figure-12}
\end{figure*}

To try to identify a set of LIME tuning parameter values with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:figure-12}). \kge{\sout{All density simulation methods performed well based on the three metrics. This is an interesting result, since Figure \ref{fig:figure-11} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend.}} \kge{The performance of a set of tuning parameter values often do not agree across metrics. For example, both density methods perform well according to average fidelity but poorly according to MSEE, and all quantile bin methods perform well accroding to average $R^2$ but poorly based on average fidelity and MSEE. However, there is consistency in results across different Gower powers.} The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME. \kge{\sout{Perhaps we can say that the kernel density method or 3 equal bins are the best performing methods, but unfortunately, Figure \ref{fig:figure-12} leaves us without a clear indication of which set of explanations to use to explain the random forest.}} \kge{It is not obvious which set of of tuning parameters are best, but perhaps, we can say that based on all three metrics, the 4 equal bins with a Gower power of 0.5 performs bettter than the other tuning parameter values considered.}

To provide a more detailed view of explanations obtained using different tuning parameter values, we take a closer look at explanations for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match), using explanation scatterplots. Figure~\ref{fig:figure-13} includes the visual representation plots of explanations (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 4 equal bins and 4 quantile bins are included to depict tuning parameter values with good (4 equal bins) and mediocre (4 quantile bins) performance based on the assessment metric plot. Plots for the kernel density simulation method are included in Appendix~\ref{exp-scatter_plus} (Figure \ref{fig:figure-B2}).

First, focus on the explanation for case NM with 4 quantile bins. \kge{The explanation plot from \emph{lime} shows that observed values with rough correlation (rough\_cor) greater than 0.05 and mismatches value less than or equal to 7.17 support a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with rough correlation greater then 0.05 and mismatches less than 7.17 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of cross correlation function (ccf) greater than 0.28 and less than or equal to 0.35 supports a non-match, and the scatter plots shows that the simulated values within this region have mostly random forest predictions less than 0.5.  However, the random forest assigns a probability of 0.19 to case NM, which is not explained by the LIME explanation. Instead, consider that a better explanation based on the scatterplots would be that the random forest assigned a prediction probably close to 0 since the observation has a rough correlation less than 0.5, mismatches greater than 4, and cross correlation function 0.6.} \kge{\sout{Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls in and supports the random forest prediction of a non-match.}}

The other explanation scatterplots can be interpreted in a similar manner. Overall, the plots identify that the \kge{\sout{3} 4} quantile bin explanations are not faithful to the random forest model predictions. The \kge{4 \sout{3}} equal bin explanations are more faithful to the random forest, but the regions captured by the bins do not align \kge{as well as they could} with the regions \kge{containing \sout{with}} similar probabilities produced by the random forest.

@@

@
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: \kge{4 \sout{3}} quantile bins (first column) and \kge{4 \sout{3}} equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}

@@

In addition to assessing faithfulness by visually comparing the complex and explainer model, we propose a visual comparison of two faithfulness metrics (MSEE and average fidelity). The examples of faithfulness metric comparisons in this paper (Figures \ref{fig:figure-07} and \ref{fig:figure-12}) both produced conflicting results. \kge{\sout{The density based simulation methods result in the best or close to the best performance even though the feature heatmaps (Figures \ref{fig:figure-06} and \ref{fig:figure-11}) show that the density simulation methods produce global explanations with minimal variation in features selected by LIME. For the bin based simulation methods, the two metrics often do not agree or contradict one another, which} This} makes it difficult to decide on a recommendation of a set of tuning parameter values that produce the explanations with the most faithful explainer model.

@@

The final claim, \textbf{localness}, is addressed by the feature heatmap and metric comparison plots. As stated, the feature heatmap reveals that the density simulation methods in the \kge{\data \ example \sout{examples of this paper}} results in global explanations where the same features are repeatedly chosen across all (or almost all) observations in the set of explanations. This finding agrees with that of \citet{laugel:2018} who found LIME produced global explanations with the normal approximation simulation method. For the bin based simulation methods in the bullet data example, the feature heatmap showed that the features chosen for the explanations varied between the two classification categories (match versus non-match). This is an interesting finding that suggests that different features can play a role in the predictions of observations in different response categories. Furthermore, while the metric comparison plot in the bullet example does not provide agreement between metrics on a best bin based method, all metrics agree that a Gower power of 0.5 for computing the model weights associated with distance of a simulated data point from the prediction of interest is the best. This suggests that a less local explanation provides a better explanation of the performance of the random forest.

@@

Section \ref{exp-scatter} introduces explanation scatterplots under the default simulation method in the \emph{lime} R package: four quantile bins. The structure of an explanation scatterplot remains the same if any bin based simulation method is used, i.e., any number of quantile or equally spaced bins. However, if the kernel density or normal approximation simulation methods are used, the format of the explanation scatterplot changes. In the density based simulation method scenarios, LIME uses the standardized versions of the predictor variables to fit the explainer model. Thus, the explainer model needs to be represented differently in the explanation scatterplot.

>>>>

We apply LIME using six sets of tuning parameter values to all observations in the \data \ test set to observe variability across tuning parameter values. A quantile bin based simulation method (samples are simulated uniformly from a specified number of quantile bins) is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, because the complexity of the explanations increases with the number of bins. Note that 4-quantile-bins are the default method in \emph{lime}. The sixth application of LIME uses a kernel density simulation method (samples are drawn from kernel density approximations of the feature distributions). The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. (See Appendix~\ref{lime-details} for descriptions of the tuning parameters.)

@@

Before introducing the visual diagnostics, let us consider a commonly used visualization of a LIME explanation that is shown in Figure \ref{fig:figure-04} (left). The plot is created using the \emph{lime} R package and provides a visual representation of a LIME explanation for the prediction of interest indicated in \autoref{fig:figure-03}. In particular, the explanation is obtained using 4-quantile-bins to simulate the data. In this scenario, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the explainer model features.

@@

Based on the plot on the left of \autoref{fig:figure-04} alone, it is not possible to make an informed assessment of the explanation. For a further assessment of the explainer model, we turn to an \emph{explanation scatterplot}: a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method of 4-quantile-bins.

@@

First, consider the claim of localness. The weights decay relatively slowly  outside of the intersection of the 2-quantil-bin suggesting that the LIME explanation is highly influenced by points outside of the bins containing the prediction of interest. However, it is  difficult to say if the claim of localness has been violated, because a definition of a local region is not specified. Depending on what region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless, it is possible to say that the weights assigned to the simulated data do not capture  the local region identified by the intersection of the quantile bins.

Now, consider the claim of faithfulness: It can be said that the majority of the points in the $x_2$ quantile support a prediction of blue, which is captured by the bar supporting a prediction of blue in the LIME explanation, and a similar statement can be made about the $x_1$-quantile-bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest performs well at capturing the sine curve classification boundary by creating various sized rectangles consisting of predictions with similar probabilities, which the LIME explanation does not pick up on. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.

@@

\autoref{fig:figure-06} shows a feature heatmap for the LIME applications to the 100 observations in the \data \ test set. The most important and second most important features selected by lime are shown in in the top and bottom facets, respectively. For the quantile bins, the original features prior to the indicator variable transformation are included since it is obvious that different features would be selected when the sizes of the bins change. This figure shows that for kernel density and 2-quantile-bins, LIME selects $x_1$ as the most important feature and $x_2$ as the second most important feature across all cases in the test set. These explanations are not local. There is variability in the features selected by LIME for 3- to 6-quantile-bins suggesting more local explanations. There are signs of vertical striping, which suggests a dependence on tuning parameters. Note that the explanations from 3- and 5-quantile-bins include the selection of the random noise variable ($x_3$) as an important variable in many predictions, which should not be the case. The pattern seen in the explanations for 3- to 6-quantile-bins may suggest local explanations, but the dependence on tuning parameters makes it unclear which set of explanations to use.

@@

All three metrics suggest that the kernel density method performed well, but the metrics disagree for the quantile bins methods. Average $R^2$ and average fidelity rank the performance of the number of quantile bins the same (2-quantile-bins perform the best and 6-quantile-bins perform the worst). In fact, these two metrics appear to have a mirrored  relationship in this example. MSEE provides almost the exact opposite results with  6-quantile-bins performing the best and 2-quantile-bins performing the worst. Average fidelity and MSEE are similar metrics, so it is not surprising that they would agree in this example. Another factor might be that MSEE only takes the prediction of interest  into account and not the full simulated dataset. The contradiction between metrics makes it difficult to identify which simulation method to trust.

@@

In response, \citet{hare:2017} proposed an automated machine learning method for bullet matching to complement a visual inspection by firearm examiners. Based on high-resolution topological scans of land engraved areas, \citet{hare:2017} obtain signatures of striations from two bullet lands (\autoref{fig:figure-09}). Nine features quantifying the similarity of signatures, such as the cross-correlation function, the distance between signatures, and the number of matching striae, are extracted and used to train a random forest to determine the probability of a comparison resulting from the same source (matching signatures) or from different sources (non-matching signatures). The model in \citet{hare:2017} was trained on a set of scans of bullets from the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}, which included 10,384 land-to-land comparisons \citet{hare:2017}. See the supporting information for additional information on the signature similarity features.

@@

\hh{XXX minor change to emphasize that this is not a completly new model, just an expanded training set - otherwise you open yourself up to criticism of not using a validated model. The bulletxtrctr model was used in an external validation in \citet{validation}. }
\hh{Here, we train a } random forest model using the same nine comparison features as \citet{hare:2017} to predict whether  \hh{two lands come from the same source} (i.e.\ are a match) in an expanded set of \Sexpr{format(dim(bullet_train)[1], big.mark = ",")} land-to-land comparisons from \hh{two sets of} bullets in the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}. We apply the trained random forest model to 6 bullets from \hh{another set of} the Hamby study with \Sexpr{dim(bullet_test)[1]} rows of land comparisons. See the supporting information for further descriptions of the data.

We first consider a global visualization of the relationship between the random forest predictions and the model features with a parallel coordinate plot of the training data (top facet) and testing data (bottom facets)  (\autoref{fig:figure-10}).  In the training data we observe a clear difference in feature values based on whether the corresponding random forest scores are close to 1 or 0. High values of rough correlation, cross correlation function, and number of matches are indicative of random forest scores close to 1; similarly, low values of mismatches, distance, and non-consecutively matching striae are also associated with random forest scores close to 1.
Observations in the test data that are classified incorrectly by the random forest tend to have feature values similar to observations in the training data with similar random forest scores. For example, the observations in the test set known to be a match but have a random forest score below 0.5 have relatively low values of rough correlation, cross correlation function, and matches and relatively large values of mismatches, distance, and non-consecutivley matching striae compared to observations with a random forest probability close to 1.

LIME is applied to all test set observations using different tuning parameter values: 12 sampling methods (2 to 6 equally spaced bins, 2 to 6 quantile-bins, kernel density estimation, and normal approximation) and 3 Gower exponents (0.5, 1, and 10). Thus, a total of $12\times 3=36$ different applications of LIME are performed. We specify that each LIME explanation return 3 features and feature selection is performed by selecting the features with the highest weights in a ridge regression model (default option).

@@

First, applications of 2-quantile-bin simulations produce the same explanations for almost all cases and LIME tuning parameter values. This suggests that the LIME explanations are global and not local. Second, within a simulation method, the features selected by LIME for an observation do not appear to vary by the Gower power. However, the LIME explanation for an observation often varies across simuluation methods. With the-equal-bins, there are vertical stripes that suggest a dependence of the LIME explanations on the number of bins. The vertical stripes are not as apparent with the quantile bins. Lastly, there are clear differences between the LIME explanations produced by the bin based simulation methods for the matches and non-matches. This suggests that different features are of importance in the random forest, depending on whether the observation corresponds to match or non-match.

@@

@
\caption{Assessment metric plot for the applications of LIME on the bullet comparison data test set. There are discrepancies in performance across metrics for a set of tuning parameter values, but based on the overall performance of all three metrics, 4-equal-bins with a Gower power of 0.5 appears to perform the best.}
\label{fig:figure-12}
\end{figure*}

To try to identify a set of LIME tuning parameter values with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:figure-12}). The performance of a set of tuning parameter values often do not agree across metrics. For example, both density methods perform well according to average fidelity but poorly according to MSEE. All quantile bin methods perform well according to average $R^2$ but poorly based on average fidelity and MSEE. However, there is consistency in results across different Gower powers. The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME.  It is not apparent which set of tuning parameters is the best. Considering all three metrics, we might conclude that the 4-equal-bins with a Gower power of 0.5 performs better than the other tuning parameter values considered.

To provide a more detailed view of explanations obtained using different tuning parameter values, we take a closer look at explanations for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match), using explanation scatterplots. Figure~\ref{fig:figure-13} includes the visual representation plots of explanations (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 4-equal-bins and 4-quantile-bins are included to depict tuning parameter values with good (4-equal-bins) and mediocre (4-quantile-bins) performance based on the assessment metric plot. Plots for the kernel density simulation method are included in Appendix~\ref{exp-scatter_plus} (Figure \ref{fig:figure-B2}).

First, focus on the explanation for case NM with 4-quantile-bins. The explanation plot from \emph{lime} shows that observed values with rough correlation (rough\_cor) greater than 0.05 and mismatches value less than or equal to 7.17 support a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with rough correlation greater than 0.05 and mismatches less than 7.17 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of cross correlation function (ccf) greater than 0.28 and less than or equal to 0.35 supports a non-match, and the scatter plots shows that the simulated values within this region have mostly random forest predictions less than 0.5.  However, the random forest assigns a probability of 0.19 to case NM, which is not explained by the LIME explanation. Instead, consider that a better explanation based on the scatterplots would be that the random forest assigned a prediction probably close to 0 since the observation has a rough correlation less than 0.5, mismatches greater than 4, and cross correlation function 0.6.

The other explanation scatterplots can be interpreted in a similar manner. Overall, the plots identify that the 4-quantile-bin explanations are not faithful to the random forest model predictions. The 4-equal-bin explanations are more faithful to the random forest, but the regions captured by the bins do not align well with the regions containing  similar probabilities produced by the random forest.

@@

@
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 4-quantile-bins (first column) and 4-equal-bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}

@@

In addition to assessing faithfulness by visually comparing the complex and explainer model, we propose a visual comparison of two faithfulness metrics (MSEE and average fidelity). The examples of faithfulness metric comparisons in this paper (Figures \ref{fig:figure-07} and \ref{fig:figure-12}) both produced conflicting results. This makes it difficult to decide on a recommendation of a set of tuning parameter values that produce the explanations with the most faithful explainer model.

@@

The final claim, \textbf{localness}, is addressed by the feature heatmap and metric comparison plots. As stated, the feature heatmap reveals that the density simulation methods in the \data \ example  results in global explanations where the same features are repeatedly chosen across all (or almost all) observations in the set of explanations. This finding agrees with that of \citet{laugel:2018} who found LIME produced global explanations with the normal approximation simulation method. For the bin based simulation methods in the bullet data example, the feature heatmap showed that the features chosen for the explanations varied between the two classification categories (match versus non-match). This is an interesting finding that suggests that different features can play a role in the predictions of observations in different response categories. Furthermore, while the metric comparison plot in the bullet example does not provide agreement between metrics on a best bin based method, all metrics agree that a Gower power of 0.5 for computing the model weights associated with distance of a simulated data point from the prediction of interest is the best. This suggests that a less local explanation provides a better explanation of the performance of the random forest.

@@

Section \ref{exp-scatter} introduces explanation scatterplots under the default simulation method in the \emph{lime} R package: 4-quantile-bins. The structure of an explanation scatterplot remains the same if any bin based simulation method is used, i.e., any number of quantile or equally spaced bins. However, if the kernel density or normal approximation simulation methods are used, the format of the explanation scatterplot changes. In the density based simulation method scenarios, LIME uses the standardized versions of the predictor variables to fit the explainer model. Thus, the explainer model needs to be represented differently in the explanation scatterplot.
